\section{Heat Equation}
The heat conduction within a medium can be described using the following partial differential equation (PDE):
\begin{gather}
\frac{\partial u}{\partial t} = \alpha \nabla^{2} u
\end{gather}
With \(u\) being a function of space and time and \(\alpha\) being a positive constant.
For this paper \(u\) will be defined in terms of one spacial dimension:
\begin{gather}
u := u(x, t) \\
\frac{\partial u}{\partial t} = \alpha \frac{\partial^{2} u}{\partial x^{2}} \\
x \in \chi \subset \mathbb{R} \quad t \in \tau \subset \mathbb{R} \\
x_{0} \leq x \leq x_{n} \quad t_{0} \leq t \leq t_{n} 
\end{gather}
\cite{Gustafsson2011}

In order to not only model the conduction of heat within a medium but also a heating process, a new function \(h: \chi \times \tau \rightarrow  \mathbb{R}\) is introduced:
\begin{gather}
\frac{\partial u}{\partial t} = \alpha \frac{\partial^{2} u}{\partial x^{2}} + h(x,t) \label{eq-1d-h}
\end{gather}
For this paper it is assumed that initial and boundary values are known:

\begin{gather}
a, b \in \mathbb{R} \\
f: \chi \rightarrow \mathbb{R} \\
u(x_{0}, t) = a \quad u(x_{n}, t) = b \\
u(x, t_{0}) = f(x) 
\end{gather} 

Applying the Fourier transform (FT) w.r.t \(x\) to \ref{eq-1d-h} yields the inhomogeneous ordinary differential equation (ODE):

\begin{gather}
\hat{u} = \mathfrak{F}(u) \quad \hat{h} = \mathfrak{F}(h) \\
\frac{d}{dt} \hat{u} = -\alpha\omega^{2}\hat{u} + \hat{h} \label{eq-1d-h-ft}
\end{gather}

A solution to \ref{eq-1d-h-ft} is given by:
\begin{gather}
\hat{u} = \hat{u}_{0} + \hat{u}_{p}
\end{gather}
Where \(\hat{u}_{0}\) is the homogeneous solution and \(\hat{u}_{p}\) is the particular integral.
In order to solve this ODE for the particular integral \(\hat{h}\) has to be known. \cite{Papula2015}
The choice of \(h\) is, except to some restrictions, arbitrary.
Therefore an approximate solution to \ref{eq-1d-h-ft} \(\hat{u}_{a}\) is obtained by the forward euler scheme:
\begin{gather}
\frac{d}{dt} \hat{u} \approx \frac{\Delta \hat{u}}{\Delta t} \\
\hat{u}_{t+1} = \hat{u}_{t} + \Delta t (-\alpha\omega^{2}\hat{u} + \hat{h}) \label{eq-1d-h-es} \\
\hat{u}_{a} = [\hat{u}_{t_{0}}, ..., \hat{u}_{t_{n}}]
\end{gather}
In order to apply the euler scheme successfully an initial condition   \(\hat{u}_{0}\) has to be known. 
This initial condition is obtained by applying the discrete Fourier transform (DFT) to an initial temperature distribution along \(x\):
\begin{gather}
\hat{u}_{0} = \mathfrak{F}(f(x)) 
\end{gather} 
\cite{Gustafsson2011b}

The forward euler scheme is used here because it is fairly easy to implement.
By applying the  inverse discrete Fourier transform (IDFT) to \(\hat{u}_{a}\) an approximate solution to \ref{eq-1d-h} can be obtained.
To decrease computing time, the fast Fourier transform (FFT) and inverse fast Fourier transform (IFFT) is used instead of the DFT and IDFT.
